G\"{o}del's Ontological Proof of God's Existence

### 2013-08-18 Natural Deduction/GodProof-ND.tex
Remove notes from draft
They were mostly based on an incorrect notion of Necessitation

\author{Bruno Woltzenlogel Paleo}

@@

\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$ G(x) $}
        \AXC{$ $} \RightLabel{2}
        \UIC{$\psi(x)$}
        \UIC{$G(x) \imp \psi(x)$}
        \UIC{$\all x.[G(x) \imp \psi(x)]$} \RightLabel{Necessitation}
        \UIC{$\nec \all x.[G(x) \imp \psi(x)]$} \RightLabel{2}
        \UIC{$\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]$}
        \UIC{$\all \psi.[\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]]$}
    \BIC{$G(x) \wedge \all \psi.[\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]]$} \dottedLine
    \UIC{$\ess{G}{x}$} \RightLabel{1}
    \UIC{$G(x) \imp \ess{G}{x}$}
    \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
\end{prooftree}

\noindent
\textbf{Note:} the formal proof above uses the necessitation rule of the basic modal logic \textbf{K}, instead of using Axiom 4, which corresponds to a restricted form of necessitation. I currently do not know how to formally prove theorem 3 using axiom 4 and not relying on necessitation.

@@

\noindent
\textbf{Note:} Theorem A could be proved more quickly using the necessitation rule. Interestingly, the proof above shows that, by using the given axioms and definitions of god and necessary existence, theorem A can be derived even without the necessitation rule.

@@

\noindent
\textbf{Note:} The proof above relies on a theorem of the modal logic \textbf{S5}, which is a quite strong modal logic. It would be interesting to try to derive theorem 4 with weaker modal axioms.

>>>>

\author{Bruno Woltzenlogel Paleo, Annika Siders}

@@

Let the following derivation with the open assumption $G(x)$ be $\Pi_1$:

\begin{prooftree}
\AXC{$ \neg P(\psi)^1$} 
       \AXC{$ $} \RightLabel{Ax. 2}
       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\forall$E}  
       \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp$E}   
 \BIC{$P(\neg\psi)$}\RightLabel{} 
           	\AXC{$G(x)$} \dottedLine\RightLabel{Definition of G}
		 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall$E}
              	\UIC{$P(\varphi)\imp \varphi(x)$}\RightLabel{$\imp$ E}
           \BIC{$ \neg \psi(x)$}\RightLabel{$\imp$E}  
            \AXC{$ \psi(x)^2$} \RightLabel{$\imp$E}  
             \BIC{$\bot $}\RightLabel{RAA, 1}  
              \UIC{$ P(\psi)$}\RightLabel{$\imp$ I, 2}
              \UIC{$ \psi(x)\imp P(\psi)$}
\end{prooftree}

Let the following derivation with the open assumption $G(x)$ be $\Pi_2$: 
\begin{prooftree}
       \AXC{$\psi(x)^1$} \RightLabel{}
             	\AXC{$\Pi_1 $}  \dashedLine
		\UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp$E}
       		\BIC{$P(\psi)$}\RightLabel{$\imp$E}  
		\AXC{$$} \RightLabel{Ax. 4}
       		\UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\forall$E}  
		\UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp$E}  
          \BIC{$\Box P(\psi)$}\RightLabel{$\imp$I, 1} 
           \UIC{$\psi(x)\imp\Box P(\psi)$} 
\end{prooftree}

Let the following derivation without open assumptions be $\Pi_3$:

\begin{prooftree}
       \AXC{$P(\psi)^1$} \RightLabel{}
             	\AXC{$G(x)^2$} \dottedLine\RightLabel{Definition of G}
		\UIC{$\forall\varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall$E}
		\UIC{$P(\psi)\imp \psi(x)$}\RightLabel{$\imp$E}  
       		\BIC{$\psi(x)$}\RightLabel{$\imp$I, 2}  
       		\UIC{$G(x)\imp \psi(x)$}\RightLabel{$\forall$I} 
		\UIC{$\forall x .(G(x)\imp \psi(x))$} \RightLabel{$\imp$I, 1}  
		\UIC{$P(\psi)\imp \forall x .(G(x)\imp \psi(x))$}
\end{prooftree}

Let the following derivation with the open assumption $G(x)$ be $\Pi_4$: 
\begin{prooftree}
       \AXC{$\psi(x)^1 $} \RightLabel{}
             	\AXC{$\Pi_2$}  \dashedLine
		\UIC{$\psi(x)\imp\Box P(\psi)$}\RightLabel{$\imp$E}
		\BIC{$\Box P(\psi)$}
		  	\AXC{$\Box P(\psi)^2 $}\RightLabel{$\Box$E}  
       			\UIC{$P(\psi)$}
				\AXC{$\Pi_3$}  \dashedLine\RightLabel{}
       				\UIC{$P(\psi)\imp \forall x .(G(x)\imp \psi(x))$}\RightLabel{$\imp$E}  
       			\BIC{$\forall x .(G(x)\imp \psi(x))$}\RightLabel{Necessitation}
			\UIC{$\Box \forall x .(G(x)\imp \psi(x))$}  \RightLabel{$\imp$I, 2} 
			\UIC{$\Box P(\psi)\imp \Box \forall x .(G(x)\imp \psi(x))$} \RightLabel{$\imp$E} 
         	 \BIC{$\Box \forall x .(G(x)\imp \psi(x)$}\RightLabel{$\imp$I, 1}
           \UIC{$\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x))$}  
\end{prooftree}

The use of the necessitation rule is valid, because the only open assumption $\Box P(\psi)$ is boxed.

We construct a derivation of theorem 3 with a subderivation $\Pi_4[G(x)^1]$, which means that the open assumption $G(x)$ in $\Pi_4$ is discharged with the rule labeled 1.

\begin{prooftree}
       \AXC{$G(x)^1 $} \RightLabel{}
             	\AXC{$\Pi_4[G(x)^1]$}  \dashedLine
		\UIC{$\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x))$}\RightLabel{$\forall $I}
		\UIC{$\forall \psi .(\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x)))$}\RightLabel{$\&$I}
		\BIC{$G(x)\& \forall \psi .(\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x)))$}\dottedLine\RightLabel{Definition of ess} 
       			\UIC{$\ess{G}{x}$} \RightLabel{$\imp$I, 1}
			\UIC{$G(x)\imp \ess{G}{x}$}
			  \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
\end{prooftree}

@@

@@
